Revealing transient strain in geodetic data with Gaussian process regression

### 2017-10-19 RevisedSubmission/hines_hetland_2017_gji.tex
did more

We make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$. Using a more concise notation, we describe each component of $\vec{u}$ as $u_i \sim \mathcal{GP}\left(0,C_{u_i}\right)$. The function $C_{u_i}$ must be positive definite, which means that the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of points $\mitbf{P}$. We assume that $C_{u_i}$ can be separated into spatial and temporal functions as 
\begin{equation}\label{eq:TransientCovariance}
C_{u_i}(p,p') = C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = \phi^2 X_i(\vec{x},\vec{x}')T_i(t,t'),
\end{equation}   
where the parameter $\phi$ controls the amplitude of realizations of $u_i$. As long as the functions $X_i$ and $T_i$ are positive definite, $C_{u_i}$ is guaranteed to also be positive definite and thus a valid covariance function \citep[sec. 4.2.4]{Rasmussen2006}. We also require that the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist everywhere, which ensures that $u_i$ is spatially and temporally differentiable (See \citet[sec. 2.2]{Adler1981} for a formal definition of differentiation for Gaussian processes). As an example, we can satisfy our requirements for positive definiteness and differentiability by using a squared exponential function for $X_i$ and $T_i$,
\begin{equation}\label{eq:ExampleCov}
X_i(\vec{x},\vec{x}') = \exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2\ell^2}\right), \ \ \ \ T_i(t,t') = \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
\end{equation}
The parameters $\ell$ and $\tau$ control the length-scale and time-scale, respectively, of realizations of $u_i$. While the squared exponential function is a commonly used covariance function for GPR, it is not appropriate for every application. The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe. To keep this section sufficiently general, we hold off on specifying $X_i$ and $T_i$ until Section \ref{sec:SignalModel}, where we estimate transient strain from slow slip events in the Pacific Northwest.

@@

respectively.

>>>>

We make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$. Using a more concise notation, we can equivalently state our prior on each component of $\vec{u}$ as $u_i \sim \mathcal{GP}\left(0,C_{u_i}\right)$. The function $C_{u_i}$ must be positive definite, which means that the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of points $\mitbf{P}$. We assume that $C_{u_i}$ can be separated into spatial and temporal functions as 
\begin{equation}\label{eq:TransientCovariance}
C_{u_i}(p,p') = C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = \phi^2 X_i(\vec{x},\vec{x}')T_i(t,t'),
\end{equation}   
where the parameter $\phi$ controls the amplitude of realizations of $u_i$. As long as the functions $X_i$ and $T_i$ are positive definite, $C_{u_i}$ is guaranteed to also be positive definite and thus a valid covariance function \citep[sec. 4.2.4]{Rasmussen2006}. We also require that the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist. This ensures that $u_i$ is spatially and temporally differentiable, allowing us to compute $\dot\varepsilon$ (See \citet[sec. 2.2]{Adler1981} or \citet[sec. 10A]{Papoulis1991} for a definition of stochastic differentiation and the conditions for differentiability).

We now provide an example to give our prior on $\vec{u}$ a more tangible meaning. We can satisfy our requirements for positive definiteness and differentiability by using a squared exponential function for $X_i$ and $T_i$,
\begin{equation}\label{eq:ExampleCov}
X_i(\vec{x},\vec{x}') = \exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2\ell^2}\right), \ \ \ \ T_i(t,t') = \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
\end{equation}
The parameters $\ell$ and $\tau$ control the length-scale and time-scale, respectively, of realizations of $u_i$. An example realization is shown in Figure ...

While the squared exponential function is a commonly used covariance function for GPR, it is not appropriate for every application. The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe. To keep this section sufficiently general, we hold off on specifying $X_i$ and $T_i$ until Section \ref{sec:SignalModel}, where we estimate transient strain from slow slip events in the Pacific Northwest.

@@

respectively \citep[sec. 10]{Papoulis1991}.
